### 2024-09-17 teaching/math240f24/draft.tex
Lecture proof techniques

Notice how we introduced the variable $x$ in the above proof.
Because we're trying to prove a ``for all'' statement, we use
the word ``let,'' to indicate that $x$ is given to us by some higher power. In fact, we should sort
of view this higher power as possibly being malicious. A very useful way to think about writing proof is
to imagine a game in which you are trying to prove a statement in predicate logic, and a supernatural adversary
is attempting to thwart you. Let's say this predicate has four variables, so the statement is
$$\exists x \, \forall y \, \forall z\, \exists w : P(x,y,z,w).$$
The variables in the statement are introduced from left to right. Each time you see a $\exists$
symbol, it's your turn. In the example above, you get to set the variable $x$ to any element of the
given universe (keeping in mind that your eventual goal is to prove $P(x,y,z,w)$). Each time there
is a $\forall$ symbol, it's the adversary's turn, and you should be prepared for whatever he throws at you.
So in our example, the adversary may set $y$ and $z$ to anything in the universe, and he knows you picked
$x$. Lastly, you get to pick $w$. If $P(x,y,z,w)$ is true, you win, and if not, the adversary wins.
Writing a proof is equivalent to describing a winning strategy for the player against the adversary.
\bye

Many mathematical statements introduce some hypotheses, then assert some conclusion. Thus they are some
kind of statement of the form $p\To q$. To prove this kind of statement,
we assume that $p$ holds, then prove that $q$ is true. This is because the only way a statement for $p\To q$
to be false is if $p$ is true and $q$ is false, so we're showing this cannot happen. Take a look at
the following example.

@@

If $n$ is odd, then $n=2k+1$ for some integer $k$, and
$$1 + (-1)^{2k+1}\bigl(2(2k+1) - 1\bigr) = 1- (4k+2-1) = 1-(4k-1) = -4k,$$
which is also a multiple of $4$.\slug

>>>>

Notice how we introduced the variable $x$ in the above proof.
Because we're trying to prove a ``for all'' statement, we use
the word ``let,'' to indicate that $x$ is given to us by some higher power. In fact, we should sort
of view this higher power as possibly being malicious. A very useful way to think about writing a proof is
to imagine a game in which you are trying to prove a statement in predicate logic, and a supernatural adversary
is attempting to thwart you. Let's say this predicate has four variables, so the statement is
$$\exists x \, \forall y \, \forall z\, \exists w : P(x,y,z,w).$$
The variables in the statement are introduced from left to right. Each time you see a $\exists$
symbol, it's your turn. In the example above, you get to set the variable $x$ to any element of the
given universe (keeping in mind that your eventual goal is to prove $P(x,y,z,w)$). Each time there
is a $\forall$ symbol, it's the adversary's turn, and you should be prepared for whatever he throws at you.
So in our example, the adversary may set $y$ and $z$ to anything in the universe, and he knows you picked
$x$. Lastly, you get to pick $w$. If $P(x,y,z,w)$ is true, you win, and if not, the adversary wins.
Writing a proof is equivalent to describing a winning strategy for the player against the adversary.

\medskip\noindent\boldlabel\datestamp{17}{ix} Proving conditional statements.
Many mathematical statements introduce some hypotheses, then assert some conclusion. Thus they are some
kind of statement of the form $p\To q$. To prove this kind of statement,
we assume that $p$ holds, then prove that $q$ is true. This is because the only way a statement for $p\To q$
to be false is if $p$ is true and $q$ is false, so we're showing this cannot happen. Take a look at
the following example.

@@

If $n$ is odd, then $n=2k+1$ for some integer $k$, and
$$1 + (-1)^{2k+1}\bigl(2(2k+1) - 1\bigr) = 1- (4k+2-1) = 1-(4k-1) = -4k,$$
which is also a multiple of $4$.\slug
\bye
